# Lola binding to opening promoters

``` {r header_child, child="/lola_paper/aws/analysis/child_docs/header_child.Rmd"}
```
``` {r setup, echo=FALSE, message=FALSE, results="hide"}
options(knitr.figure_dir = "lola_binding_to_opening_promoters")
library(GenomicRanges)
library(plyr)
library(magrittr)
library(ggplot2)
library(reshape2)
library(DESeq2)
library(doMC)
library(dplyr)

source("/lola_paper/aws/analysis/shared/knitr_common.r")
source("/lola_paper/aws/analysis/shared/granges_common.r")
source("/lola_paper/aws/analysis/shared/flybase.r")
source("/lola_paper/aws/analysis/shared/tf_motif_enrichments.r")

#pcutoff=0.01
````

```{r libraries, include=FALSE} 



GRangesFromDataFrames<-function(dataframe){with(dataframe,GRanges(IRanges(start=start,end=end),seqnames=seqnames,strand=strand,dataframe%>%dplyr::select(-strand,-start,-end,-seqnames,-width)))}

````

````{r}
get_lola_peaks<-function(location="promoter",number=1000,promoter_width=500,upstream=50){

import("/lola_paper/aws//bed/lola_i_peaks_1_summits.bed") %>% resize(fix='center',width=200) ->lola_peaks_1.gr   
import("/lola_paper/aws//bed/lola_i_peaks_2_summits.bed") %>% resize(fix='center',width=200) ->lola_peaks_2.gr
lola_peaks_1.gr%<>%{.[order(.$score,decreasing=TRUE)][1:number]}
lola_peaks_2.gr%<>%{.[order(.$score,decreasing=TRUE)][1:number]}

lola_peaks.gr <- lola_peaks_1.gr[findOverlaps(lola_peaks_1.gr,lola_peaks_2.gr) %>% queryHits %>% unique()] %>% resize(fix='center',width=1)
valid.chrs<-c("chr2L","chr2R","chr3L","chr3R","chrX","chr4")
lola_peaks.gr%<>%keepSeqlevels(valid.chrs,pruning.mode="coarse")
lola_peaks.gr%<>%assign_nearest_gene(.)
lola_peaks.gr$location<-'distal'
lola_peaks.gr$location[flybase_txs_granges() %>% resize(fix='start',width=upstream) %>% resize(fix='end',width=promoter_width-upstream) %>% findOverlaps(lola_peaks.gr) %>% subjectHits]<-'promoter'
lola_peaks.gr%<>%.[.$location %in% location]
lola_peaks.gr
}

```{r load_gene_groups_and_further_sub_group, include=TRUE, echo=TRUE, message=TRUE}


comparisson_groups.lst <-readRDS("/lola_paper/aws/analysis/figure1b_enrichments/comparisson_groups.lst.rds")


txs.gr<-flybase_txs_granges()

late.gr <- txs.gr[txs.gr$fb_tx_id %in% (comparisson_groups.lst$late.df)$fb_tx_id] %>% resize(fix="start",width=1) %>% resize(fix="end",width=200)

nochange.gr <- txs.gr[txs.gr$fb_tx_id %in% (comparisson_groups.lst$nochange.df)$fb_tx_id] %>% resize(fix="start",width=1) %>% resize(fix="end",width=200)

groups <- readRDS("/lola_paper/aws/analysis/figure1e_lola_targets_heatmap/groups.rds")

sum(late.gr$fb_gene_id %in% txs.gr$fb_gene_id[txs.gr$fb_tx_id %in% groups$lola_promoters])

sum(late.gr$fb_gene_id %in% {get_lola_peaks(number = 750)}$nearest_gene_id)
length({get_lola_peaks(number = 750)}$nearest_gene_id)
sum(late.gr$fb_gene_id %in% {get_lola_peaks(number = 1000)}$nearest_gene_id)
length({get_lola_peaks(number = 1000)}$nearest_gene_id)
sum(late.gr$fb_gene_id %in% {get_lola_peaks(number = 1500)}$nearest_gene_id)
length(late.gr$fb_gene_id)
length({get_lola_peaks(number = 1500)}$nearest_gene_id)

sum(nochange.gr$fb_gene_id %in% {get_lola_peaks(number = 750)}$nearest_gene_id)
sum(nochange.gr$fb_gene_id %in% {get_lola_peaks(number = 1000)}$nearest_gene_id)
sum(nochange.gr$fb_gene_id %in% {get_lola_peaks(number = 1500)}$nearest_gene_id)

sum(nochange.gr$fb_gene_id %in% txs.gr$fb_gene_id[txs.gr$fb_tx_id %in% groups$lola_promoters])

length(nochange.gr$fb_gene_id)
````

````
``` {r session_info, echo=FALSE, comment=NA}
sessionInfo()
```
