``` {r setup, echo=FALSE, message=FALSE, include=FALSE, error=FALSE}
library(cowplot)
library(dplyr)
library(tidyr)
library(magrittr)
library(pander)
library(xtable)
library(ggplot2)
library(parallel)
library(reshape2)
library(matrixStats)
library(GOstats)
library(org.Dm.eg.db)
library(dplyr)
# Output folder for this document
options(knitr.figure_dir = "figure1b_enrichments")

source("shared/knitr_common.r")
source("shared/granges_common.r")
source("shared/flybase.r")
source("shared/promoter_element_enrichments.r")
cores_to_use<-32;
path_to_data_1<-"../bw";

GRangesFromDataFrames<-function(dataframe){with(dataframe,GRanges(IRanges(start=start,end=end),seqnames=seqnames,strand=strand,dataframe%>%dplyr::select(-strand,-start,-end,-seqnames,-width)))}


```

# Calculating Pol II enrichments

``` {r header_child, child="child_docs/header_child.Rmd"}
```

## Samples overview

We will calculate the pol II enrichments for the following samples
``` {r samples_table}
samples.df <- read.csv("data/figure1b_polII_data.csv", stringsAsFactors=FALSE)
samples.df %>% pander
```

## load the samples

``` {r functions, include=FALSE}

locate_bigwig <- function(name) {
    path1 <- file.path(path_to_data_1, name)
    if(file.exists(path1)) return(path1)
  path2 <- file.path(path_to_data_2, name)
  if(file.exists(path2)) return(path2)
  stop("Cannot find bigwig: ", name)
}

enrichments_for_sample <- function(i, samples.df, txs.gr) {
  sample.bw <- locate_bigwig(samples.df$ip[i])
  wce.bw    <- locate_bigwig(samples.df$wce[i])

  label <- samples.df$label[i]
  factor_type <- samples.df$factor[i]
  
  regions.gr <- promoters(txs.gr, upstream=samples.df$window_upstream[i], downstream=samples.df$window_downstream[i])
  
  mcols(txs.gr)$region_ip  <- regionSums(regions.gr, sample.bw)
  mcols(txs.gr)$region_wce <- regionSums(regions.gr, wce.bw)

  mcols(txs.gr)$ip_total <- total_signal(sample.bw)
  mcols(txs.gr)$wce_total <- total_signal(wce.bw)

  mcols(txs.gr)$region_wce_min_coverage <- regionMins(regions.gr, wce.bw)

  mcols(txs.gr)$label <- label
  mcols(txs.gr)$factor <- factor_type

  txs.gr
}

txs.gr <- flybase_txs_granges() %>% filter_chrs

  data.df <- 1:nrow(samples.df) %>%
                 mclapply(enrichments_for_sample, samples.df, txs.gr, mc.cores=cores_to_use, mc.preschedule=FALSE) %>%
                 lapply(as.data.frame) %>%
                 bind_rows 
  data.df

```

## Calculate enrichments

Before calculating enrichment, we floor the WCE signal for each region at the median WCE signal level among all transcripts.

``` {r calculate_enrichments, include=FALSE}

enrichments.df <- data.df %>% 
                      group_by(label, factor) %>%
                      mutate(median_region_wce = median(region_wce))

enrichments.df %<>% transform(enrichment = (region_ip / ip_total) / (pmax(region_wce, median_region_wce) / wce_total))

tss.infs <- which(is.infinite(enrichments.df$enrichment))

min.tss <- with(enrichments.df, min(enrichment[is.finite(enrichment)]))

if(length(tss.infs) > 0) enrichments.df$enrichment[tss.infs] <- min.tss

```

## Save results

``` {r save_results, include=FALSE}
enrichments.df %<>%
  ungroup %>%
  dplyr::select(fb_tx_id, fb_gene_id, fb_symbol, enrichment, factor, label)
  
#filter for highest tx_id
enrichments.df %<>% filter(fb_tx_id %in% (enrichments.df %>% group_by(fb_gene_id) %>% summarize(fb_tx_id = fb_tx_id[which.max(enrichment)]) %>% `$`(fb_tx_id)))

enrichments.df %>% saveRDS(file=figure_path("enrichments.df.rds"))

```

#different groups

``` {r load_data, include=FALSE}

supp.df<-read.csv(file="data/bjoern_groups.csv", header = TRUE, sep = ",");

flybase_ids.df<-read.csv(file="data/FlyBase_IDs.txt", header = TRUE, sep = "\t");

flybase_ids.df %>% mutate(flybase_gene_id=submitted_id,fb_gene_id=converted_id) %>% merge(supp.df,.) ->supp.df

supp.df<-supp.df[!supp.df$fb_gene_id %>% duplicated,]

supp.df%<>% merge(enrichments.df %>% filter(label == "early_1") %>% dplyr::select(fb_gene_id,fb_tx_id))

all_promoters.df = enrichments.df %>% filter(label == "early_1") %>% dplyr::select(fb_gene_id,fb_tx_id)

supp.df %>% filter(Poised_constant_set==1) -> nochange.df
		 
supp.df %>% filter(Poised_opening_set==1)  -> late.df

comparisson_groups.lst=list(all_promoters.df=all_promoters.df,late.df=late.df,nochange.df=nochange.df)				 
saveRDS(comparisson_groups.lst, file=figure_path("comparisson_groups.lst.rds"))

```

## Promoter element heatmap

``` {r build_heatmap, include=FALSE}

gene_groups <- list("Late"=late.df$fb_tx_id,
                    "NoChange"=nochange.df$fb_tx_id)

names(gene_groups) <- names(gene_groups) %>%
                      paste0("\n", pn(elementNROWS(gene_groups)), " genes")

results.df <- promoter_element_enrichments(gene_groups)

g <- promoter_element_heatmap(results.df, "", manual.scale=c(-4, 4))
```

``` {r display_heatmap, fig.cap="", dev=c("png","pdf"),fig.width=10, fig.height=3}
theme_set(theme_cowplot())
g
```

## GO terms

``` {r calc_go_enrichments, include=FALSE}

flybase_to_eg <- function(fb_ids) {
  na.omit(unlist(mget(as.character(fb_ids), org.Dm.egFLYBASE2EG, ifnotfound=NA)))
}

get_gene_names_for_bpid <- function(bpid, go_results) {
  matching <- geneIdsByCategory(go_results, bpid)[[1]]
  paste(sort(unlist(na.omit(mget(matching, org.Dm.egSYMBOL, ifnotfound=NA)))), collapse=", ")
}

run_go <- function(fb_ids, universe_fb_ids) {
  test_ids <- flybase_to_eg(fb_ids)
  universe_ids <- flybase_to_eg(universe_fb_ids)
  
  go.params <- new("GOHyperGParams", 
                   geneIds=test_ids,
                   universeGeneIds=universe_ids,
                   annotation="org.Dm.eg.db",
                   ontology="BP",
                   pvalueCutoff=0.01,
                   conditional=TRUE,
                   testDirection="over")
  go.results <- hyperGTest(go.params)

  results.df <- summary(go.results)
  results.df$Gene_Names <- sapply(results.df$GOBPID, get_gene_names_for_bpid, go.results)  
  results.df
}

go_results.late <- run_go(late.df$fb_gene_id, all_promoters.df$fb_gene_id)


go_results.nochange <- run_go(nochange.df$fb_gene_id, all_promoters.df$fb_gene_id)

```

### Late genes

See full spreadsheet in folder ``r figure_path()``

``` {r go_results_late_table, results="asis"}
write.table(go_results.late, file=figure_path("go_results_late_genes.xls"), sep="\t", row.names=F, col.names=T, quote=F)
go_results.late$Gene_Names <- NULL
head(subset(go_results.late, Size > 5), 15)
```


### No change genes

See full spreadsheet in folder

``` {r go_results_nondiff_table, results="asis"}
write.table(go_results.nochange, file=figure_path("go_results_nochange_genes.xls"), sep="\t", row.names=F, col.names=T, quote=F)
go_results.nochange$Gene_Names <- NULL
head(subset(go_results.nochange, Size > 5), 15)
```
## Normalize for heatmap

```{r Normalize_data}

promoter.df<-rbind(late.df,nochange.df)

enrichments.df %<>% spread(key=label, value=enrichment)	%>%
	 mutate(early=rowMeans(cbind(early_1,early_2)),late=rowMeans(cbind(late_1,late_2,late_3))) %>%
	  dplyr::select(-early_1,-early_2,-late_1,-late_2,-late_3) %>%
	   gather(label,enrichment,-1:-4)

enrichments.df[enrichments.df<2]<-0
enrichments.df %<>% group_by(label, factor) %>%
            mutate(normalized_enrichment = pmax(0, pmin(1, enrichment / quantile(enrichment, 0.95))))

enrichments.df %<>% filter(fb_tx_id %in% promoter.df$fb_tx_id)	


		
enrichments.df %<>% filter(label %in% c("early","late"))

saveRDS(enrichments.df, file=figure_path("polII_by_group.df.rds"))

```

## Pol II heatmap early late

```{r polII_heatmap, dev=c("png","pdf"), fig.width=14, fig.height=10}

enrichments.df %>% as.data.frame %>% mutate(label=paste0(label,".",factor)) %>% 
			dplyr::select(fb_tx_id,label,normalized_enrichment) %>% 
			spread(key= label, value = normalized_enrichment) %>% 
			dplyr::select(fb_tx_id,early.polii,late.polii) -> values.df

build_heatmap_plot <- function(groups) {
  heatmap.df <- groups %>%
                lapply(function(tx_ids) {
                  subset(values.df, fb_tx_id %in% tx_ids)
                }) %>%
                bind_rows(.id="group_name") %>% gather(time,enrichment,-group_name,-fb_tx_id) 


  heatmap.df$time%<>% factor(levels=c("early.polii","late.polii"))
  
  heatmap.df$group_name%<>% factor(levels=c("Opening","Constant"))
  heatmap.df$fb_tx_id%<>% factor(levels=heatmap.df$fb_tx_id %>% unique)
  heatmap.df%<>%{.[order(.$fb_tx_id),]}
  
  g <- ggplot(heatmap.df, aes(x=time, y=fb_tx_id, fill=enrichment)) +
       geom_tile() +
       scale_fill_gradient("Pol II enrichment", high="#3d5782", low="white") +
       facet_grid(group_name ~ ., scales="free_y", space="free_y") +
       theme_bw() +
       labs(x="", y="") +
       scale_x_discrete(expand=c(0, 0)) +
       scale_y_discrete(expand=c(0, 0)) +
       theme(axis.text.y=element_blank(),
             axis.ticks=element_blank(),
             panel.grid.major=element_blank(),
             panel.grid.minor=element_blank(),
             axis.text.x=element_text(angle=45, vjust=1, hjust=1))
  
  list(plot=g, data=heatmap.df)
}

heatmap_all   <- list(Opening = late.df$fb_tx_id, Constant = nochange.df$fb_tx_id) %>% build_heatmap_plot
heatmap_all$plot
saveRDS(heatmap_all, file=figure_path("heatmap_all.lst.rds"))


```

``` {r lola_motif, dev=c("png","pdf")}
txs.gr <- flybase_txs_granges()

txs.gr %>% as.data.frame %>% filter(fb_tx_id %in% nochange.df$fb_tx_id) %>% GRangesFromDataFrames -> nochange.gr
nochange.gr$change<-"Constant"
txs.gr %>% as.data.frame %>% filter(fb_tx_id %in% late.df$fb_tx_id) %>% GRangesFromDataFrames -> late.gr
late.gr$change<-"Opening"
promoter.gr<-c(nochange.gr,late.gr)
promoter.gr %<>% resize(fix="start",width=100) %>% resize(fix="end",width=300)
lola_g_match<-vmatchPattern(pattern="AAAGCTY",subject=Dmelanogaster,max.mismatch=0,fixed=FALSE)

mcols(promoter.gr)$lola_motif<-countOverlaps(promoter.gr,lola_g_match,ignore.strand=TRUE)
#promoter.gr$lola_motif[promoter.gr$lola_motif>1]<-1


 promoter.df<-promoter.gr %>% as.data.frame

pairwise.wilcox.test(promoter.df$lola_motif,promoter.df$change,p.adjust='hochberg')	 
			 
    build_heatmap_plot <- function(groups,promoter.df) {

      heatmap.df <- groups %>%
   						     lapply(function(tx_ids) {
   						                   promoter.df %>%
   						                     filter(fb_tx_id %in% tx_ids) %>%
   						                     dplyr::select(fb_tx_id, lola_motif) %>% 
   											 gather(motifs, Number_of_motifs, -fb_tx_id)
   						                 }) %>%
   						      bind_rows(.id="group_name")
  
  

heatmap_order<-heatmap_all %>% `$`(data)
heatmap.df$fb_tx_id%<>% factor(levels=heatmap_order$fb_tx_id %>% unique)
heatmap.df$group_name%<>% factor(levels=c("Opening","Constant"))

          g <- ggplot(heatmap.df, aes(x=motifs, y=fb_tx_id, fill=Number_of_motifs)) +
       geom_tile() +
       scale_fill_gradient("Motif enrichment", high="red", low="white") +
       facet_grid(group_name ~ ., scales="free_y", space="free_y") +
       theme_bw() +
       labs(x="", y="") +
       scale_x_discrete(expand=c(0, 0)) +
       scale_y_discrete(expand=c(0, 0)) +
       theme(axis.text.y=element_blank(),
             axis.ticks=element_blank(),
             panel.grid.major=element_blank(),
             panel.grid.minor=element_blank(),
             axis.text.x=element_text(angle=45, vjust=1, hjust=1))

list(plots=g, data=heatmap.df)
}
		 
			heatmap   <- split(promoter.df$fb_tx_id, promoter.df$change) %>% build_heatmap_plot(promoter.df)

	  	heatmap$plots




````

``` {r session_info_child, child="child_docs/session_info_child.Rmd"}
```
